### 2013-12-03 paper/main.tex
created good examples of scaled/shifted model
The description of the engine response with a scaled/shifted
has been added.  The response should serve as a good basis
for building controllers and comparing the results.

\begin{figure}[hbp!]
\begin{center}

@@

\end{center}
\caption{Linear engine model of a modern 4.6L V-8.}\label{fig:lem}
\end{figure}

@@

% {{{ fig:em_plot
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.\protect\footnotemark}
\label{fig:em_plot}
\end{figure}
\footnotetext{The Matlab source to produced these plots is in Appendix \ref{app:em_matlab}.}
% }}}

@@

Applying this equation results in the response shown
in Figure \ref{fig:rpmtime_plot}.
It can be seen it takes over two seconds for the system to
stabilize with no control.
Clearly this is unacceptably slow and performance could be
improved by using a controller.

% {{{ fig:rpmtime_plot
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_rpmtime_plot}
\end{center}
\caption{Engine rpm response compared to time with a constant
0.65 input control.\protect\footnotemark}
\label{fig:rpmtime_plot}
\end{figure}
\footnotetext{The Matlab source of the RPM vs time plot is in Appendix \ref{app:rpmtime_plot_source}.}
% }}}

@@

\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}
\label{app:rpmtime_plot_source}

@@

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of engine model
		versus time.}
	]{../octave/engine_model_rpmtime_plot.m}

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of the normalized engine model.}
	]{../octave/norm_engine_model_plot.m}

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of the normalized engine model
	after step is applied.}
	]{../octave/norm_engine_plot_close.m}

>>>>

% {{{ fig:lem
\begin{figure}[hbp!]
\begin{center}

@@

\end{center}
\caption{Linear engine model of a modern 4.6L V-8.}\label{fig:lem}
\end{figure}
% }}}

@@

% {{{ fig:em_plot
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.\protect\footnotemark}
\label{fig:em_plot}
\end{figure}
\footnotetext{The Matlab source for this and the other engine model plots
are in Appendix \ref{app:em_matlab}.}
% }}}

@@

Applying this equation results in the response shown
in Figure \ref{fig:rpmtime_plot}.
It can be seen that the rise time over $0.5$ second and
that the overshoot is close to $40\%$.
The stabilization time is also several seconds.
Clearly its performance could be improved with the application
of a controller.

% {{{ fig:rpmtime_plot
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_rpmtime_plot}
\end{center}
\caption{Engine rpm response compared to time with a constant
0.65 input control.}
\label{fig:rpmtime_plot}
\end{figure}
% }}}

The response from zero rpm is not a realistic measurement of performance
because zero rpm is impossibly slow.
A typical idle rpm range is from 600 to 1000 rpm.

Figure \ref{fig:em_shift_plot2} shows the response for a step input
with control inputs chosen to reach 600 and 1000 rpm
\footnote{The control input to stable rpm values are given in
Appendix \ref{app:contrpm}}.
This is a more realistic representation of typical operation.
The rise time is near $0.2$ seconds and the overshoot is $15\%$.
And the stabilization time is almost one second.
Its behavior can still be improved with a controller.

% {{{ fig:em_shift_plot2
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.6]{../octave/em_shift_plot2}
\end{center}
\caption{Response for a shifted and scaled step input with zero torque.}
\label{fig:em_shift_plot2}
\end{figure}
% }}}

@@

% {{{ Engine Model Matlab Source
\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}
\label{app:rpmtime_plot_source}

@@

% XXX
\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of engine model
		versus time.}
	]{../octave/engine_model_rpmtime_plot.m}

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of engine model
		versus time for a shifted/scaled step input.}
	]{../octave/em_shift_plot2.m}
% }}}

\clearpage
\section{Control Response}
\label{app:contrpm}

For a given control input with zero torque the stable rpm values reached
are shown in Figure \ref{fig:stable}.
Equation \ref{eq:stable} describes its behavior.
Listing \ref{lst:stable} shows how this was calculated.

\begin{align}
y &= 1450.1 x \label{eq:stable}
\end{align}

\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.5]{../octave/stable_rpm_plot}
\end{center}
\caption{Stable rpm reached for a given constant control input.}
\label{fig:stable}
\end{figure}

\clearpage
\lstinputlisting[
	caption={Matlab source to find the stable rpm for a given
		control input with zero torque.},
	label={lst:stable}
	]{../octave/stable_rpm_plot.m}
